Forecasting biodiversity in breeding birds using best practices

### 2017-12-08 manuscript/draft.Rmd

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. While each model could explain most of the variance
in richness, none of them consistently outperformed a baseline model that
predicted constant richness at each site. In particular, we found no evidence
that current methods (such as species distribution models) can successfully turn
spatial data into useful temporal predictions about biodiversity at decadal
time-scales. Most methods' prediction intervals were also too narrow: twice as
many observations fell outside the predicted 95% interval as should be expected.
The intervals produced by "stacking" species distribution models---perhaps the
most common forecasting approach for terrestrial species richness---were
especially misleading, failing to contain the true value at 4-6 times the
nominal rate.  The best practices implemented in this study directly influence
the forecasts, the relative performance of different modeling approaches, and
the conclusions about the current state of biodiversity forecasting. To
facilitate the rapid improvement of biodiversity forecasts, we emphasize the
value of specific best practices in making forecasts and evaluating forecasting
methods.

>>>>

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. All models explained more than 50% of the variance in
richness, but none of them consistently outperformed a baseline model that
predicted constant richness at each site. The best practices implemented in this
study directly influenced the forecasts and evaluations. Stacked species
distribution models and "naive" forecasts produced poor estimates of uncertainty
and accounting for this resulted in these models dropping in the relative
performance compared to other models. Accounting for observer effects improved
model performance overall, but also changed the rank ordering of models because
it did not improve the accuracy of the "naive" model. Considering the forecast
horizon revealed that the prediction accuracy decreased across all models as the
time horizon of the forecast increased. To facilitate the rapid improvement of
biodiversity forecasts, we emphasize the value of specific best practices in
making forecasts and evaluating forecasting methods.
